Modelling last glacial cycle ice dynamics in the Alps

### 2018-07-21 articles/cycle/response-referee2.tex

Thank you very much for your detailed review of our manuscript.

% ----------------------------------------------------------------------

\sechead{General comments}

    \referee{%
        This manuscript describes a model study with the Parallel Ice Sheet
        Model applied to the last glacier cycle in the Alps. The climate
        forcing is derived from present day climate of WorldClim and the
        ERA-Interim reanalysis and time-dependent temperature offsets derived
        from the Greenland Ice core (GRIP), from Antarctic ice core (EPICA) and
        Marine sediment core from the Iberian margin (MD01-2444). The study is
        split in two, in the first part analysis of six model simulations (with
        and without precipitation scaling) made on a 2 km resolution grid is
        presented and concluded that out of the three climate forcing records
        used, the EPICA record gives the most realistic ice volume history
        during MIS 4 and 2. The second part analyses simulation made with the
        EPICA forcing record on a 1km grid. The authors draw conclusions about
        the ice cover, ice flow pattern, ice thickness, LGM ice extent, which
        in their model is a transient stage with varying timing across their
        model domain due to glacier dynamics.}

    Thank you very much for this accurate summary.

    \referee{%
        The manuscript is well organized
        and clearly written, the missing thing in this study is a discussion of
        and preferably a sensitivity study of the ice dynamic-model assumptions
        made. Is the sliding of the ice age ice sheet realistically modelled
        with pseudo-plastic assumption using the Shallow Shelf approximation?
        How sensitive are the results to the selected model parameters? It is
        briefly mentioned once on page 19, line 8, but a thorough analysis of
        the model sensitivity would strengthen the paper.}

    We agree that a sensitivity study to ice physics, in particular basal
    sliding, is one of many improvements needed upon our results. However, for
    the kind of exercise presented here, the sensitivity of to palaeo-climate
    history, which is virtually unknown and has many degrees of freedom, far
    outweight the model sensitivity to ice physics. For this manuscript we have
    chosen to focus on the primary source of sensitivity, which is climate.

    Because the kind of simulations presented here depends on computing
    resources only available through a peer-review application process, we are
    currently unable to perform an additional sensitivity study on basal
    sliding for the entire Alps.  Regional sentivity studies are more
    accessible. A regional sensitivity analysis on basal sliding is precisely
    the topic of a follow-up study currently under preparation for the Rhine
    Glacier \citep{Imhof.etal.2017}, which has a relatively isolated catchment
    for its size as compared to other Alpine Glaciers.

    Meanwhile, the following text was added in the section on trimlines, which
    is certainly the mostly sensitive aspect of our results to basal sliding:

    \msquote{%
        The remaining discrepancies may relate to temporal migrations of the
        basal thermal boundary, an absence of sliding in warm-based areas, and
        levelling of small-scale topographic features in the 1\,km horizontal
        grid. They call for more detailed comparisons spanning the entire
        Alpine range and specific sensitivity studies to relevant basal sliding
        and ice rheological parameters, and to the uncertain subglacial
        topography.}

    Basal sliding and other sources of uncertainties in ice physics where also
    mentioned in the conclusions:

    \msquote{%
        However, these results are limited by uncertainties on ice physics. The
        till deformation model used here does not hold for sliding over bedrock
        surfaces. On the other hand, the constant friction angle used is
        representative of wet till but weaker basal conditions may have applied
        over saturated lake sediments where they occured. A constant rheology
        was used for temperature ice containing more than 1\,\% of liquid
        water.}

### 2018-07-21 articles/cycle/alps-cycle.tex
Address referee 2 comment on defining age scale.

    From the mid-twentieth
    century, palaeo-climate records extracted from deep sea sediments and
    ice cores have provided a much more detailed picture of the Earth's
    environmental history \citep[e.g.,][]{Emiliani.1955,
    Shackleton.Opdyke.1973, Dansgaard.etal.1993, Augustin.etal.2004},
    indicating several tens of glacial and interglacial periods. During the
    last 800\,ka before present, these \emph{glacial cycles}
    have succeeded each other with a periodicity about 100\,ka \citep{Hays.etal.1976,
    Augustin.etal.2004}. Nevertheless, this global signal is largely governed
    by the North American and Eurasian ice sheet complexes. It is thus unclear
    whether glacier advances and retreats in the Alps were in pace with global
    sea-level fluctuations.
    Besides, the landform record is typically sparse in time and, most often,
    spatially incomplete. Palaeo-ice sheets did not leave a continuous imprint
    on the landscape, and much of this evidence has been overprinted by
    subsequent glacier re-advances and other geomorphological processes
    \citep[e.g.,][]{Kleman.1994, Kleman.etal.2006, Kleman.etal.2010}. Dating
    uncertainties typically increase with age, such that dated reconstructions
    are strongly biased towards later glaciations \citep{Heyman.etal.2011}.
    In the European Alps, although sparse
    geological traces indicate that the last glacial cycle may have comprised
    two or three cycles of glacier growth and decay \citep{Preusser.2004,
    Ivy-Ochs.etal.2008}, most glacial features currently left on the foreland
    present a record of the last major glaciation of the Alps, dating from the
    \emph{Last Glacial Maximum} \citep[LGM;][]{Ivy-Ochs.2015,
    Wirsig.etal.2016, Monegato.etal.2017}.

>>>>

    From the mid-twentieth
    century, palaeo-climate records extracted from deep sea sediments and
    ice cores have provided a much more detailed picture of the Earth's
    environmental history \citep[e.g.,][]{Emiliani.1955,
    Shackleton.Opdyke.1973, Dansgaard.etal.1993, Augustin.etal.2004},
    indicating several tens of glacial and interglacial periods. During the
    last 800\,ka (thousand years before present), these \emph{glacial cycles}
    have succeeded each other with a periodicity about 100\,ka \citep{Hays.etal.1976,
    Augustin.etal.2004}. Nevertheless, this global signal is largely governed
    by the North American and Eurasian ice sheet complexes. It is thus unclear
    whether glacier advances and retreats in the Alps were in pace with global
    sea-level fluctuations.
    Besides, the landform record is typically sparse in time and, most often,
    spatially incomplete. Palaeo-ice sheets did not leave a continuous imprint
    on the landscape, and much of this evidence has been overprinted by
    subsequent glacier re-advances and other geomorphological processes
    \citep[e.g.,][]{Kleman.1994, Kleman.etal.2006, Kleman.etal.2010}. Dating
    uncertainties typically increase with age, such that dated reconstructions
    are strongly biased towards later glaciations \citep{Heyman.etal.2011}.
    In the European Alps, although sparse
    geological traces indicate that the last glacial cycle may have comprised
    two or three cycles of glacier growth and decay \citep{Preusser.2004,
    Ivy-Ochs.etal.2008}, most glacial features currently left on the foreland
    present a record of the last major glaciation of the Alps, dating from the
    \emph{Last Glacial Maximum} \citep[LGM;][]{Ivy-Ochs.2015,
    Wirsig.etal.2016, Monegato.etal.2017}.
